A Recurrence

### 2014-12-10 draft.tex
project done

\newtheorem{proposition}{Proposition}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}

@@

In this section we consider the case where $F(x,y) = f(y)/x$ with $f(y) = y^n$, where $n$ is any complex number, not just an integer. The question we would like to answer is given an integer $p$ can we always come up with an $n$ such that $F$ has period $p$. We first explore some basic properties of the sequences such functions produce and compute an explicit example. We then give an explicit formula for the $k$th term in the sequence. Finally using a generating function we are able to produce a necessary condition on $n$, in order for $F$ to produce a sequence with period $p$. Using Mathematica we were able to verify numerically for values of $p$ up to 1000 that this necessary condition is also sufficient. Because of this we conjecture 
\begin{conjecture}
\label{cos}
Given any integer $p>2$ let 
\begin{equation*}
n = 2 \cos(2 \pi l/p),
\end{equation*}
with $l \in \{1,2,3, \ldots, p-1 \}$. Then $F(x,y) = y^n/x$ is periodic with period $p$.
\end{conjecture}

@@

We note further that there is a relation between different terms of the two sequences.  
\begin{lemma} 
\label{ab} For $k\geq 0$
\begin{equation*}
\alpha_{k+1} = - \beta_{k}.
\end{equation*}
\end{lemma}
\begin{proof}
We will show this using induction. As a base case note that $\alpha_1 = 0 = \beta_0$, and $\alpha_2 = -n = \beta_1$. Now if we assume this holds for $k \leq m$ and would like to show it holds for $k = m+1$ consider $\alpha_{m+1}$ which by the recurrence relation is
\begin{equation*}
\alpha_{m+1} = n \alpha_{m} - \alpha_{m-1}.
\end{equation*}
Well $\alpha_m = -\beta_{m-1}$ and $\alpha_{m-1} = - \beta_{m-2}$ by the induction hypothesis and so 
\begin{equation*}
\alpha_{m+1} = - n \beta_{m-1} + \beta_{m-2} = -\beta_{m}.
\end{equation*}
\end{proof}
Note that this Lemma also reduces the number of simultaneous equations that we need to solve from four to three as we always have $\alpha_{p+1} = - \beta_{p}$. Both are set to zero so it is enough to only include one of them in the system of equations

We now given an exact formula for $\beta_k$ 
\begin{proposition} 
For $k \geq 1$
%= \binom{k-1}{k-1} n^{k-1} - \binom{k-2}{k-3}n^{k-3} +\binom{k-3}{k-5}n^{k-5} -\binom{k-4}{k-7}n^{k-7} + \dots \\
\begin{align*}
\beta_k =\sum_{i = 0}^{\lfloor (k-1)/2 \rfloor} (-1)^i\binom{k-1-i}{k-1-2i}n^{k-1-2i}
\end{align*}
\end{proposition}
\begin{proof}
It is easy to check that this holds for $k = 1$ and $k = 2$.  If we plug the formula into the recurrence relation
\begin{equation*}
\beta_k = n\beta_{k-1} - \beta_{k-2},
\end{equation*}
we must have 
\begin{align*}
\sum_{i = 0}^{\lfloor (k-1)/2 \rfloor} (-1)^i\binom{k-1-i}{k-1-2i}n^{k-1-2i} = & n \sum_{i = 0}^{\lfloor (k-2)/2 \rfloor} (-1)^i\binom{k-2-i}{k-2-2i}n^{k-2-2i} \\
& +\sum_{i = 0}^{\lfloor (k-3)/2 \rfloor} (-1)^i\binom{k-3-i}{k-3-2i}n^{k-3-2i}\\
\end{align*}
Checking this equation separately when $k$ is even or odd, as well as using combinatorial identities, shows that this expression for $\beta_k$ satisfies the recurrence relation. Because of this our formula is valid for $k \geq 1$.
\end{proof}

This expression for $\beta_k$, gives us an expression for $\alpha_k$ by Lemma \ref{ab}. In order to obtain a function $F$ with period $p$ we must solve the three equations
\begin{align*}
\alpha_p &=1 \\
\beta_p &=0\\ 
\beta_{p+1} &= 1,
\end{align*}
plugging in the explicit values for $\alpha_p, \beta_p$ and $\beta_{p+1}$. We were not able to show that these equations always produce a solution for $n$.
\\

Now using a generating function approach, given a desired period $p$, we give a necessary condition for $n$. 
First we produce the generating function for $\alpha_k$. A generating function of an infinite sequence is a formal power series with coefficients equal to the terms of the sequence.  So if $G(\xi)$ is the generating function of $\alpha_k$, then 
\begin{equation*}
G(\xi) = \sum_{k=0}^{\infty} \alpha_k \xi^k,
\end{equation*}
where $\xi$ is some formal variable. We can find a closed form expression for $G(\xi)$.

@@

\section{Haikus} 
Its periodic  \\
What came before is what's next \\
Its periodic \\ \\
No matter the $p$ \\
You can find a few good $n$\\
That is what we hope \\

>>>>

\newtheorem{proposition}{Proposition}
\newtheorem{definition}{Definition}
\newtheorem{lemma}{Lemma}
\newtheorem{conjecture}{Conjecture}

@@

In this section we consider the case where $F(x,y) = f(y)/x$ with $f(y) = y^n$, where $n$ is any complex number, not just an integer. The question we would like to answer is given an integer $p$ can we always come up with an $n$ such that $F$ has period $p$. We first explore some basic properties of the sequences such functions produce and compute an explicit example. We then give an explicit formula for the $k$th term in the sequence. Finally using a generating function we are able to produce a necessary condition on $n$, in order for $F$ to produce a sequence with period $p$. Using Mathematica we were able to verify numerically for values of $p$ up to 1000 that this necessary condition is also sufficient. Because of this we conjecture

\begin{conjecture}
\label{cos}
Given any integer $p>2$ let 
\begin{equation*}
n = 2 \cos(2 \pi l/p),
\end{equation*}
with $l \in \{1,2,3, \ldots, p-1 \}$. Then $F(x,y) = y^n/x$ is periodic with period $p$.
\end{conjecture}

@@

We note further that there is a relation between different terms of the two sequences.  
\begin{lemma} 
\label{ab} For $k\geq 0$
\begin{equation*}
\alpha_{k+1} = - \beta_{k}.
\end{equation*}
\end{lemma}
\begin{proof}
We will show this using induction. As a base case note that $\alpha_1 = 0 = \beta_0$, and $\alpha_2 = -n = \beta_1$. Now we assume this holds for $k \leq m$. To show it holds for $k = m+1$, consider
\begin{equation*}
\alpha_{m+1} = n \alpha_{m} - \alpha_{m-1}.
\end{equation*}
We have $\alpha_m = -\beta_{m-1}$ and $\alpha_{m-1} = - \beta_{m-2}$ by the induction hypothesis and so 
\begin{equation*}
\alpha_{m+1} = - n \beta_{m-1} + \beta_{m-2} = -\beta_{m}.
\end{equation*}
\end{proof}
Note that this Lemma also reduces the number of simultaneous equations that we need to solve from four to three as we always have $\alpha_{p+1} = - \beta_{p}$. Both are set to zero so it is enough to only include one of them in the system of equations

We now given an exact formula for $\beta_k$ 
\begin{proposition} 
For $k \geq 1$
%= \binom{k-1}{k-1} n^{k-1} - \binom{k-2}{k-3}n^{k-3} +\binom{k-3}{k-5}n^{k-5} -\binom{k-4}{k-7}n^{k-7} + \dots \\
\begin{align*}
\beta_k =\sum_{i = 0}^{\lfloor (k-1)/2 \rfloor} (-1)^i\binom{k-1-i}{k-1-2i}n^{k-1-2i}
\end{align*}
\end{proposition}
\begin{proof}
It is easy to check that this holds for $k = 1$ and $k = 2$.  If we plug the formula into the recurrence relation
\begin{equation*}
\beta_k = n\beta_{k-1} - \beta_{k-2},
\end{equation*}
we must have 
\begin{align*}
\sum_{i = 0}^{\lfloor (k-1)/2 \rfloor} (-1)^i\binom{k-1-i}{k-1-2i}n^{k-1-2i} = & n \sum_{i = 0}^{\lfloor (k-2)/2 \rfloor} (-1)^i\binom{k-2-i}{k-2-2i}n^{k-2-2i} \\
& +\sum_{i = 0}^{\lfloor (k-3)/2 \rfloor} (-1)^i\binom{k-3-i}{k-3-2i}n^{k-3-2i}.\\
\end{align*}
We omit the steps of explicitly checking the equation, but simple combinatorial identities show that both sides are equal. Thus, our formula is valid for $k \geq 1$.
\end{proof}

This expression for $\beta_k$ gives us an expression for $\alpha_k$ by Lemma \ref{ab}. In order to obtain a function $F$ with period $p$ we must solve the three equations
\begin{align*}
\alpha_p &=1 \\
\beta_p &=0\\ 
\beta_{p+1} &= 1,
\end{align*}
for $n$. We were not able to show that these equations always produce a solution, despite the fact that we know the explicit form of these equations for any $p$.
\\

Now we switch our focus and use a generating function approach to this problem. We show a necessary condition for $n$ so that $y^n/x$ produces a sequence with a given $p$.

First we produce the generating function for $\alpha_k$. Note that we shift from looking at $\beta_k$ to looking at $\alpha_k$ simply to make the generating function formulation simpler. However, we know from Lemma~\ref{ab} that these two sequences are almost interchangeable.

A generating function of an infinite sequence is a formal power series with coefficients equal to the terms of the sequence.  So if $G(\xi)$ is the generating function of $\alpha_k$, then 
\begin{equation*}
G(\xi) = \sum_{k=0}^{\infty} \alpha_k \xi^k,
\end{equation*}
where $\xi$ is some formal variable. We can find a closed form expression for $G(\xi)$.

@@

\section{Haikus} 
\emph{Perry's Haikus}\\

\noindent Its periodic  \\
What came before is what's next \\
Its periodic \\

\noindent No matter the $p$ \\
You can find a few good $n$\\
That is what we hope \\

\noindent\emph{Fermi's Haikus}\\

\noindent Try small cases first.\\
Two, three, and four seem simple.\\
Five and six less so.\\

\noindent Restrict attention\\
To the $y$ to the $n$ case.\\
Now it always works!\\

\noindent We can prove this claim.\\
Just write the equations down.\\
Shoot, how does this work?\\
